Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies

### 2024-06-16 paper.tex
small things

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Singly Vibrationally Enhanced (SIVE) spectroscopy, a type of nonlinear hyper-Raman spectroscopy, is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
We derive selection rules and coherent interference effects in singly resonant SIVE (SR-SIVE).
The theory of SR-SIVE is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant SIVE (DR-SIVE), akin to a resonant hyper-Raman process. 
DR-SIVE is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.
The linenarrowing abilities of DR-SIVE are discussed.
We showcase the difference in selection rules of SR-SIVE and DR-SIVE through a study of cyanocobalamin, highlighting the sensitivity of these measurements on non-resonant and resonant hyper-Raman hyperpolarizabilities, respectively.
\end{abstract}

@@

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
Singly vibrationally enhanced (SIVE) is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
Non-parametric, singly resonant SIVE (SR-SIVE) spectroscopy was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
In SR-SIVE, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to induce a scattering process to provide output.
SIVE was documented long ago to have characteristics similar to spontaneous hyper-Raman scattering spectroscopy. \cite{RN352}
Hyper-Raman scattering, first theorized by Decius \textit{et al.} and experimentally observed by Terhune \textit{et al.}, is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965}
Compared to Raman spectroscopy, Hyper-Raman scattering cross sections are often small, making it a difficult technique to successfully implement.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering, however, make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems.
In 1998, Cho et al. proposed a parametric six wave mixing technique which was the coherent analogue of hyper-Raman scattering. \cite{Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes,\cite{RN243, Cho2000_Cascade} it is preferable to investigate coherent four wave mixing analogues to hyper-Raman scattering.

@@

We investigate properties of doubly resonant SIVE (DR-SIVE) to understand how coupling between vibrational and vibronic states can be measured through DR-SIVE.
DR-SIVE is shown to demand one and two-photon absorptions between the ground and excited state.
DR-SIVE is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}
Based upon these selection rules, we investigate SR-SIVE and DR-SIVE spectroscopy of cyanocobalamin and find good agreement between the theoretical and experimental results.

\section{Experimental Methods}
The ultrafast laser system has been described in detail elsewhere. \cite{RN278, Kaufman2024}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG \"DF2\", Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 were computer controlled through communication with 50 mm DC actuators (PMC). %todo: who tf makes PMC motors?
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X.%todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused into the sample using a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale. %@dan if we should change to amp lmk

@@

Note that we choose the convention where an arbitrary electric field, assuming a plane wave, is written in Fourier components as 
\begin{equation}
	\vec{E}(\omega) = \frac{1}{2} \vec{E}_0 \exp(i(\vec{k}\cdot\vec{x} - \omega t)) + c.c.
\end{equation}
where $E_0$ is the pulse amplitude and $\vec{k}$ is a wavevector.

Expressions for $\gamma_{ijkl}$ can be found by solving Liouville-von Neumann equation for an output density matrix element $\rho_{ij}$.
Generally, under the electric dipole approximation, the density matrix, $\bar{\rho}(t)$, is propagated through the Liouville-von Neumann equation as \cite{RN455}
\begin{equation}\label{matrix}
	\dot{\bar{\rho}}(t) = \frac{1}{i \hbar}[H_0 - \vec{\mu}\cdot \sum_j E_j(\vec{x},t), \bar{\rho}(t)] + \bar{\Gamma} \bar{\rho}(t)
\end{equation} %check if bar or double bar
where $H_0$ is the system (time independent) Hamiltonian, $\vec{\mu}$ is the electric transition dipole moment, $\vec{E}(\vec{x},t)$ is a time dependent electric field, and $\bar{\Gamma}$ is the pure dephasing rate vector. 
In the rotating frame, $\tilde{\rho}_{ij} = \rho_{ij} e^{-i \omega_{ij}t}$.
By solving \autoref{matrix} under steady state ($\dot{\bar{\rho}}(t) = 0$) in the rotating frame, expressions for $\gamma_{ijkl}$ can be readily obtained by multiplying the output dipole moment, $\mu_i$ by the final density matrix element $\rho_{ij}$.

@@

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the total angular momentum and $M$ is its projection onto the space fixed quantization axis).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, the matrix elements in \autoref{sivebeta} can be evaluated through the Wigner-Eckart theorem.
It is important to note that $\beta_{ijk}$ does not have complete tensor index symmetry.\cite{Andrews1978}
Specifically, it has shown that the rank one spherical tensor ($T_{q_1}^{(1)}$, with $q_1 \in \{-1, 0 ,1\}$) fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)} + T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Andrews1990}
%there is some fancy way to note that theres no index constraint but i dont remember the language...
As such, its decomposition into irreducible tensor form is allowed only if the values of $q_2, q_3$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

\subsection{Quantitative SR-SIVE}
With the selection rules of SR-SIVE understood for vibrational and rovibrational spectroscopy, it becomes possible to extract quantitative information from SIVE spectra to compare to incoherent methods and other CMDS methods.
Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
In the method introduced by Levenson and Bloembergen (Bloembergen Interferometry Experiment), an internal standard interferes with the resonant lineshape.
The internal standard is often chosen to be C$_6$H$_6$ or C$_6$D$_6$, due to the strong Raman activity and measured $\chi^{(3)}$ of its ring breathing mode ($\nu_2$ in Herzberg notation), causes interference. \cite{Levenson1974_2, RN351, RN345}
The self-heterodyning method measures $\chi^{(3)}$ and does not require measurement of absolute intensities; it instead relies on self-heterodyning of the internal standard signal. 
Resonant lineshapes are also complicated by amplitude level interference between the sample and its substrate and/or sample cell windows. \cite{RN362, RN418}

Most quantitative methods in an n$^{th}$ order CMDS experiment are used to extract $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.
To our knowledge, only vibraitonal SFG and 2D-IR spectroscopy has taken advantage of orientationally averaged $\chi^{(n)}$ values to extract transition moments. \cite{Shen90, Moilanen2009, RN245}
Previous treatments did not consider using quantitative analysis in SIVE to measure $\beta_{ijk}$ values from SIVE spectra.
It is thus useful to investigate how a simple treatment of orientational averaging can extract $\beta_{ijk}$ from a $\chi^{(3)}_{IJKL}$ expression.

@@

The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. 
Infrared spectra can provide values for different $\vec{\mu}$ components.
Since $\vec{\mu}$ is a rank-one tensor, its three Cartesian components share the simple relationship $\langle {\mu^2_x} \rangle = \langle {\mu^2_y} \rangle = \langle {\mu^2_z} \rangle = \frac{1}{3}\langle {\mu^2} \rangle$. \cite{RN459}
Neglecting rotational dependence (i.e., assuming only vibrational transitions), $\langle {\mu^2} \rangle = \abs{\mu}^2 = \mu^2$, so that $\langle \beta_{ijk} \mu_l \rangle = \langle \beta_{ijk} \rangle \frac{\mu}{\sqrt{3}}$.
By introducing the reduced mass coordinate $q_n = \sqrt{\frac{1}{m_n}} Q_n$ and using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_n}$ found earlier, we see
\begin{equation}\label{betasive}
	\langle \frac{\partial \beta_{ijk}}{\partial q_n} \rangle = \varepsilon_0 \frac{48 \sqrt{3} }{NF}  \frac{\Gamma_{gv} \omega_{vg}}{\rho_{gg}} \frac{\Im(\chi^{(3)}_{IJKL})}{\frac{\partial \mu_{l}}{\partial q_n}}
\end{equation}
Since $\mu$ values can be readily extracted from FT-IR spectra,\cite{RN412} this shows that SIVE can give quantitative information on the magnitude of $\beta_{ijk}$.

A key result of \autoref{betasive} is that the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$ and $\frac{\partial \mu_{l}}{\partial q_n}$ can be extracted from SIVE spectra. 
These ideas can prove useful when assessing the validity of computational methods which report on the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$.
Similar ideas were first introduced by Shen to extract the sign of $\frac{\partial \mu_{l}}{\partial q_n}$ from vSFG spectra.\cite{Shen90}
SIVE and vSFG together provide a unique picture of extracting signs for the microscopic observables that drive nonlinear output.

@@

It is clear for c = -1 that there are four poles: $\pm i \sigma, \Delta_{ga}(0), \Delta_{ba}(0)$. 
Following \autoref{fig:contours}, the contour that wraps the upper half of the complex plane has only one pole, $i \sigma$; as such, we choose to evaluate \autoref{integraly} in the upper half of the complex plane for the case of anti-correlated modes through the residue theorem as
	\begin{equation}
		\begin{split}
			\rho_{ba} = \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \rho_{ba}(\xi) &= 2\pi i \sum_k \lim_{\xi \rightarrow \xi_k} P(\xi) \rho_{ba}(\xi) (\xi - \xi_k)\\
			&= 2\pi i \frac{\eta \sigma}{\pi} \lim_{\xi \rightarrow i\sigma} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga}(0) - \xi} \frac{1}{\Delta_{ba}(0) - \xi} (\xi - i \sigma)\\
			&= 2 \eta \sigma i \lim_{\xi \rightarrow i\sigma} \frac{1}{\xi + i\sigma} \frac{1}{\Delta_{ga}(0) - \xi} \frac{1}{\Delta_{ba}(0) - \xi}\\
			&= 2\eta \sigma i \frac{1}{2i\sigma} \frac{1}{\Delta_{ga}(0) - i\sigma} \frac{1}{\Delta_{ba}(0) - i\sigma}\\
			&= \frac{\eta}{\Delta_{ga}(i\sigma)\Delta_{ba}(i \sigma)}\\
		\end{split}
	\end{equation}

In words, we see the inhomogeneous envelope width $\sigma$ is added to the linewidth of each resonance (e.g., $\Gamma_{ag} \rightarrow \Gamma_{ag} + \sigma$).
Therefore, pathway I-a cannot linenarrow anti-correlated modes. 
\subsubsection{Correlated Modes}
For correlated modes, the poles are situated in the complex plane as shown in \autoref{fig:contours}.
It is clear that no matter which contour is chosen, there is a contribution to the final integral from either $\Delta_{ba}(0)$ or $\Delta_{ga}(0)$. 
For the below calculation, we choose to evaluate along the contour in the bottom half of the complex plane, i.e., the poles ($\xi_k$) are $-i\sigma, \Delta_{ga}$.
We stress that since the integral is evaluated over $\mathbb{R}$ and there are no degenerate poles, the results below can be obtained by evaluating the contour which wraps around the poles $i\sigma, \Delta_{ba}$ using the Residue theorem as

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Singly Vibrationally Enhanced (SIVE) spectroscopy, a type of nonlinear hyper-Raman spectroscopy, is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
%alternate name is hyper-(S/D)FG... which is more interesting and descriptive than SIVE imo. lmk your thoughts
We derive selection rules and coherent interference effects in singly resonant SIVE (SR-SIVE).
The theory of SR-SIVE is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant SIVE (DR-SIVE), akin to a resonant hyper-Raman process. 
DR-SIVE is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.
The linenarrowing abilities of DR-SIVE are discussed.
We showcase the difference in selection rules of SR-SIVE and DR-SIVE through a study of cyanocobalamin, highlighting the sensitivity of these measurements on non-resonant and resonant hyper-Raman hyperpolarizabilities, respectively.
\end{abstract}

@@

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
Singly vibrationally enhanced (SIVE) is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
Non-parametric, singly resonant SIVE (SR-SIVE) spectroscopy was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
In SR-SIVE, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to induce a scattering process to provide output.
SIVE was documented long ago to have characteristics similar to spontaneous hyper-Raman scattering spectroscopy. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965}
Compared to Raman spectroscopy, hyper-Raman scattering cross sections are often small, making it a difficult technique to successfully implement.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering, however, make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems.
In 1998, Cho et al. proposed a parametric six wave mixing technique which was the coherent analogue of hyper-Raman scattering. \cite{Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes,\cite{RN243, Cho2000_Cascade} it is preferable to investigate coherent four wave mixing analogues to hyper-Raman scattering.

@@

We investigate properties of doubly resonant SIVE (DR-SIVE) to understand how coupling between vibrational and vibronic states can be measured through DR-SIVE.
DR-SIVE is shown to demand one and two-photon absorptions between the ground and excited state.
DR-SIVE is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}
Based upon these selection rules, we investigate SR-SIVE and DR-SIVE spectroscopy of cyanocobalamin, highlighting the applicability of these methods to non-model systems.

\section{Experimental Methods}
The ultrafast laser system has been described in detail elsewhere. \cite{RN278, Kaufman2024}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG \"DF2\", Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 were computer controlled through communication with 50 mm DC actuators (PMC). %todo: who tf makes PMC motors?
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X.%todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused into the sample in a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale. %@dan if we should change to amp lmk

@@

Note that we choose the convention where an arbitrary electric field, assuming a plane wave, is written in Fourier components as 
\begin{equation}
	\vec{E}(\omega) = \frac{1}{2} \vec{E}_0 \exp(i(\vec{k}\cdot\vec{x} - \omega t)) + c.c.
\end{equation}
where $E_0$ is the pulse amplitude and $\vec{k}$ is a wavevector.
%maybe talk about maxwell wave eqn for e fields ?
Expressions for $\gamma_{ijkl}$ can be found by solving Liouville-von Neumann equation for an output density matrix element $\rho_{ij}$.
Generally, under the electric dipole approximation, the density matrix, $\bar{\rho}(t)$, is propagated through the Liouville-von Neumann equation as \cite{RN455}
\begin{equation}\label{matrix}
	\dot{\bar{\rho}}(t) = \frac{1}{i \hbar}[H_0 - \vec{\mu}\cdot \sum_j E_j(\vec{x},t), \bar{\rho}(t)] + \bar{\Gamma} \bar{\rho}(t)
\end{equation} %check if bar or double bar
where $H_0$ is the system (time independent) Hamiltonian, $\vec{\mu}$ is the electric transition dipole moment, $\vec{E}(\vec{x},t)$ is a time dependent electric field, and the elements of $\bar{\Gamma}$, $\Gamma_{ij}$, are dephasing rates. 
In the rotating frame, $\tilde{\rho}_{ij} = \rho_{ij} e^{-i \omega_{ij}t}$.
By solving \autoref{matrix} under steady state ($\dot{\bar{\rho}}(t) = 0$) in the rotating frame, expressions for $\gamma_{ijkl}$ can be readily obtained by multiplying the output dipole moment, $\mu_i$ by the final density matrix element $\tilde{\rho}_{ij}$.

@@

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the total angular momentum and $M$ is its projection onto the space fixed quantization axis).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, the matrix elements in \autoref{sivebeta} can be evaluated through the Wigner-Eckart theorem.
It is important to note that $\beta_{ijk}$ does not have complete tensor index symmetry.\cite{Andrews1978}
Specifically, it has shown that the rank one spherical tensor ($T_{q_1}^{(1)}$, with $q_1 \in \{-1, 0 ,1\}$) fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)} + T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Altmann1982, Andrews1990}
%there is some fancy way to note that theres no index constraint but i dont remember the language...
As such, its decomposition into irreducible tensor form is allowed only if the values of $q_2, q_3$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

\subsection{Quantitative SR-SIVE}
With the selection rules of SR-SIVE understood for vibrational and rovibrational spectroscopy, it becomes possible to extract quantitative information from SIVE spectra to compare to incoherent methods and other CMDS methods.
Lineshape analysis is essential for extracting quantitative information from CMDS spectra.
Scanning across resonances create dispersive lineshapes, i.e, self-heterodyning, which inform on $\Re(\chi^{(3)})$ and $\Im(\chi^{(3)})$.\cite{Levenson1974_1, Levenson1974_2}
In the method introduced by Levenson and Bloembergen (Bloembergen Interferometry Experiment), an internal standard interferes with the resonant lineshape.
The internal standard is often chosen to be C$_6$H$_6$ or C$_6$D$_6$, due to the strong Raman activity and measured $\chi^{(3)}$ of its ring breathing mode ($\nu_2$ in Herzberg notation). \cite{Levenson1974_2, RN351, RN345}
Self-heterodyning of the internal standard signal measures $\chi^{(3)}$ and does not require measurement of absolute intensities. 
Resonant lineshapes are also complicated by amplitude level interference between the sample and its substrate and/or sample cell windows, which must be accounted for to obtain quantitatively correct $\chi^{(3)}$ values. \cite{RN362, RN418}

Most quantitative methods in an n$^{th}$ order CMDS experiment are used to extract $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.
To our knowledge, only vibraitonal SFG and 2D-IR spectroscopy has taken advantage of orientationally averaged $\chi^{(n)}$ values to extract transition moments. \cite{Shen90, Moilanen2009, RN245}
Previous treatments did not consider using quantitative analysis in SIVE to measure $\beta_{ijk}$ values from SIVE spectra.
It is thus useful to investigate how a simple treatment of orientational averaging can extract $\beta_{ijk}$ from $\chi^{(3)}_{IJKL}$.

@@

The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. 
Infrared spectra can provide values for different $\vec{\mu}$ components.
Since $\vec{\mu}$ is a rank-one tensor, its three Cartesian components share the simple relationship $\langle {\mu^2_x} \rangle = \langle {\mu^2_y} \rangle = \langle {\mu^2_z} \rangle = \frac{1}{3}\langle {\mu^2} \rangle$. \cite{RN459}
Neglecting rotational dependence (i.e., assuming only vibrational transitions), $\langle {\mu^2} \rangle = \abs{\mu}^2 = \mu^2$, so that $\langle \beta_{ijk} \mu_l \rangle = \langle \beta_{ijk} \rangle \frac{\mu}{\sqrt{3}}$.
By introducing the reduced mass coordinate $q_n = \sqrt{\frac{1}{m_n}} Q_n$ and using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_n}$ found earlier, we see
\begin{equation}\label{betasive}
	\langle \frac{\partial \beta_{ijk}}{\partial q_n} \rangle = \varepsilon_0 \frac{48 \sqrt{3} }{NF}  \frac{\Gamma_{gv} \omega_{vg}}{\rho_{gg}} \frac{\Im(\chi^{(3)}_{IJKL})}{\frac{\partial \mu}{\partial q_n}}
\end{equation}
Since $\abs{\mu}$ values can be readily extracted from FT-IR spectra,\cite{RN412} this shows that SIVE can give quantitative information on the magnitude of $\beta_{ijk}$.

A key result of \autoref{betasive} is that the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$ and $\frac{\partial \mu_{l}}{\partial q_n}$ can be extracted from SIVE spectra. 
Since IR and spontaneous hyper-Raman spectroscopy report on $\Im(\chi^{(1)}) \sim \abs{\mu_{l}}^2$ and $\Im(\chi^{(5)}) \sim \abs{\beta_{ijk}}^2$,\cite{RN515} respectively, the absolute signs of $\frac{\partial \mu_{l}}{\partial q_n}$ and $\frac{\partial \beta_{ijk}}{\partial q_n}$ cannot be extracted from IR or hyper-Raman spectroscopy.
These ideas can prove useful when assessing the validity of computational methods which report on the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$.
Similar ideas were first introduced by Shen to extract the sign of $\frac{\partial \mu_{l}}{\partial q_n}$ from vSFG spectra.\cite{Shen90}
SIVE and vSFG together provide a unique picture of extracting signs for the microscopic observables that drive nonlinear output.

@@

It is clear for c = -1 that there are four poles: $\pm i \sigma, \Delta_{ga}(0), \Delta_{ba}(0)$. 
Following \autoref{fig:contours}, the contour that wraps the upper half of the complex plane has only one pole, $i \sigma$; as such, we choose to evaluate \autoref{integraly} in the upper half of the complex plane for the case of anti-correlated modes through the residue theorem as
	\begin{equation}
		\begin{split}
			\rho_{ba} = \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \rho_{ba}(\xi) &= 2\pi i \sum_k \lim_{\xi \rightarrow \xi_k} P(\xi) \rho_{ba}(\xi) (\xi - \xi_k)\\
			&= 2\pi i \frac{\eta \sigma}{\pi} \lim_{\xi \rightarrow i\sigma} \frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga}(0) - \xi} \frac{1}{\Delta_{ba}(0) - \xi} (\xi - i \sigma)\\
			&= 2\eta \sigma i \frac{1}{2i\sigma} \frac{1}{\Delta_{ga}(0) - i\sigma} \frac{1}{\Delta_{ba}(0) - i\sigma}\\
			&= \frac{\eta}{\Delta_{ga}(i\sigma)\Delta_{ba}(i \sigma)}\\
		\end{split}
	\end{equation}

In words, we see the inhomogeneous envelope width $\sigma$ is added to the linewidth of each resonance (e.g., $\Gamma_{ba} \rightarrow \Gamma_{ba} + \sigma$).
Therefore, pathway I-a cannot linenarrow anti-correlated modes. 
\subsubsection{Correlated Modes}
For correlated modes, the poles are situated in the complex plane as shown in \autoref{fig:contours}.
It is clear that no matter which contour is chosen, there is a contribution to the final integral from either $\Delta_{ba}(0)$ or $\Delta_{ga}(0)$. 
For the below calculation, we choose to evaluate along the contour in the bottom half of the complex plane, i.e., the poles ($\xi_k$) are $-i\sigma, \Delta_{ga}$.
We stress that since the integral is evaluated over $\mathbb{R}$ and there are no degenerate poles, the results below can be obtained by evaluating the contour which wraps around the poles $i\sigma, \Delta_{ba}$ using the Residue theorem as
